\paperTitle

### 2014-05-16 doc/icml2014/directed.tex
major structural changes up to bidependent set

% DONE: too wordy
%\section{Consistent Parameter Estimation in Directed Models}
\section{Directed models and exclusive views}
\label{sec:directed}

%\todo{we are using clique to refer to only hidden cliques; be explicit about using "hidden"}

% DONE: need more setup of what the goal is
%The tensor factorization method attacks the heart of the non-convexity
  %in latent variable models.
  In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
  In general directed models, the conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
which we were able to recover from {\TensorFactorize}, are not the underlying parameters.
  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
  %conditional probability tables for every clique for a class of directed
  %graphical models.
  Let us first consider an example (Section~\ref{sec:directedExample});
  Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
For ease of exposition we make the following simplifications to our presentation:
(i) we describe our algorithm entirely in the context of directed
  graphical models, though it generalizes to undirected models;
(ii) we derive our algorithm assuming infinite data; in \sectionref{sampleComplexity} we 
  describe error bounds for the finite sample regime; and
(iii) we present the algorithm solely in terms of linear algebraic operations;
  \sectionref{piecewise} provides a statistically
  more efficient estimator using composite likelihood.

\subsection{Example: directed grid model}
\label{sec:directedExample}

%\begin{figure}
%  \centering
%  \input{figures/grid-outline.tikz}
%  \caption{A directed grid model.}
%  \label{fig:grid}
%\end{figure}

%Consider the directed grid model shown in \figureref{grid} which
Consider the directed grid model shown in \figureref{approach}.
  % DONE: all the different dependency => too strong
  %captures the local dependency structures possible in
  %a Bayesian network.
The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  hidden variables $h_1, \ldots, h_4$.
The parameters of this model are the conditional probability tables
$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$. 
%Assume $O$ and $T$ have full column rank. % PL: if we put this in, we have to say something about V, which is too complex.

\paragraph{Estimating $O$}
Note that the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$; we can therefore use
  $\TensorFactorize$ from \sectionref{setup} to recover $O$.

\paragraph{Estimating $\pi$}
The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
  $\pi$ by a linear system: $\mO_1 = O \pi$. 
% PL: weird to refer to refer to this assumption from previous section
% where it's talking about another model.
%By \assumptionref{full-rank}, $O$ has full column rank and thus can be
  %inverted to recover $\pi$: $\pi = O\pinv  \mO_1$.
If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.

@@

%Let us now generalize the intuitions of the directed grid to general graphs.
The intuitions of the directed grid generalize readily to general graphs.
% PL: this prose is redundant with property
%Note that in each of the above cases,
  %for any clique $\sC$, each hidden variable $h_i \in \sC$ has its own
  %view which is conditionally independent given the other hidden variables in
  %the clique.
%Fundamentally, algorithm takes advantage of having a view for every
  %hidden variable in a clique to help identify its hidden state.
The key property required for our algorithm to work is as follows:
%We capture this general property as follows:
\begin{property}[Exclusive views]
  \label{prop:exclusive-views}
A hidden clique $\sC \in \sG$ is said to possess the \emph{exclusive views property} if for
  every hidden variable $h_i \in \sC$, (i)
  there exists some observed variable,
  $x_{v}$ which is conditionally independent of the rest of the clique
  given $h_i$, and (ii) the conditional moments $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ can be estimated consistently.
We call $x_v$ an exclusive view of $h_i$ in $\sC$.
If every clique has the exclusive views property, then we say that $\sG$ does as well.
\end{property}

Next, we state the non-degeneracy assumptions under which our general
  algorithm is guaranteed to obtain consistent estimates.
\begin{assumption}[Non-degeneracy]
  \label{asm:non-degeneracy}
  (i) The marginal distribution of each hidden variable $h_i$ have
  full support: $\mPi{i} \succ 0$. 
  (ii) The conditional moments $\mOpp{v}{i}$ have full column rank $k$.
\end{assumption}

Intuitively, we need to ensure that every marginal distribution
  $\Pr(x_v)$ can be separated into $k$ different conditional distributions
  $\Pr(x_v | h_i)$. 
The full rank assumption asks that the conditional distributions are
  linearly independent; if this were not the case, the conditional
  distribution for a particular latent state could be imitated by the
  combination of several other latent states.

% PL: to save space
%This property also extends to any connected subgraph $\sG' \in \sG$.
%For the rest of this section, we will only consider subgraphs which are
  %cliques. 
%In \sectionref{undirected}, we will show how the generalization
  %to subgraphs allows us to efficiently estimate the pseudo-likelihood.

% \begin{property}(Exclusive views)
%   \label{prop:exclusive-views}
%   A hidden clique $\sC$ of $\sG$ is said to possess the \textbf{exclusive views
%   property} if for every hidden variable $h_i \in \sC$,
%   there is some observed variable, $x_{v}$, which is conditionally
%   independent of the rest of the clique given $h_i$,
%   and whose conditional
%   moment $\mOpp{v}{i} \eqdef \Pr(x_{v} \mid h_i)$ can be estimated consistently.
% We call $x_v$ the exclusive view of $h_i$ in $\sC$.
% \end{property}

@@

We now show this condition is sufficient to learn the marginal
  distribution of the clique.
Consider any hidden clique $\sC = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be any exclusive view for $h_{i_j}$ in $\sC$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $\sC$.
We can write down the moments $\mO_\sV$ as follows:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  %&= \sum_{\vh \in \sH}
      %\Pr(\Sh{\sC}) \Pr(\Sx{\sV} \given \Sh{\sC}) \\
      &= \sum_{\vh_\sC} \Pr(\Sh{\sC}) 
          \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
    &= Z_{\sC}(\mOpp{v_1}{i_1},\cdots,\mOpp{v_m}{i_m}).
\end{align*}

If each $\mOpp{v_j}{i_j}$ has full column rank, then we can recover the
hidden marginals $Z_\sC$ by inverting:
\begin{align*}
  Z_{\sC} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}
\algorithmref{learnclique} summarizes the procedure, \LearnClique.
Given $Z_\sC$, the conditional probability tables for $\sC$ can easily be
obtained via renormalization (see \sectionref{sampleComplexity}).

\begin{algorithm}
  \caption{\LearnClique~(pseudoinverse)}
  \label{algo:learnclique}
  \begin{algorithmic}
    \REQUIRE Hidden clique $\sC$ with exclusive views (\propertyref{exclusive-views}).
    \ENSURE Marginal distribution of the clique $Z_\sC$.
      \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
      \STATE Return $\mH_\sC \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
  \end{algorithmic}
\end{algorithm}
% PL: sample complexity, not rank
%Of course, for $Z_{\sC}$ to be identifiable, we need various rank conditions,
%which will be established in \sectionref{sampleComplexity}.

@@

% DONE: transition the content more gradually (checkpoint: what have we done so far, what next)
% DONE: refer to properties by name in addition to by number
We have shown how to estimate the parameters of any clique that possesses the exclusive
  views property (\propertyref{exclusive-views}).
  But in a general graph, which cliques have this property?
  To provide intuition, we will provide more interpretable sufficient conditions
  and examples.

%In the example above, we were able to guarantee \propertyref{exclusive-views}
  %by identifying a hidden variable, $h_1$ along with three conditionally
  %independent observed variables, $x^a_1, x^b_1, x^a_2$ using the tensor
  %factorization method.
  %followed by using \TensorFactorize to learn the conditional moments.
First, let us revisit the definition of a bottleneck:
\begin{definition}[Bottleneck]
  A hidden variable $h_i$ is said to be a \textbf{bottleneck} if there
  exists at least three observed variables (views), $x_{v_1}, x_{v_2}, x_{v_3}$
  that are conditionally independent given $h_i$,
  and $\mOpp{v_1}{i},\mOpp{v_2}{i},\mOpp{v_3}{i}$ have full column rank $k$.
  % PL: need to add rank constraint; otherwise, doesn't imply exclusive views
  Let $\sV_{h_i}$ denote any set of views for $h_i$.
\end{definition}

% DONE: don't talk about parameter recovery yet, just establish exclusive views
%Finally, the claim we make is that we can recover parameters for any
  %graphical model where every hidden variable is a bottleneck:
Now define the following property:
\begin{property}[Uniformly bottlenecked]
  \label{prop:bottleneck}
  A clique $\sC$ is \emph{uniformly bottlenecked} if
  every hidden variable in $\sC$ is a bottleneck. 
  If every clique $\sC \in \sG$ is uniformly bottlenecked, then we say
  that the graph $\sG$ is uniformly bottlenecked as well.
\end{property}
Note that two hidden variables are allowed to have overlapping views.
%If every hidden variable is a bottleneck, then we can estimate $\mOpp{v}{i}$
It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well.

To rule out edge cases, we assume the graphical model only has conditional
independences given by the graph:
\begin{property}[Faithful representation]
  \label{prop:ci}  
  %If there is an active trail from any two variables $a, b$ in the graph
    %$\sG$, then they are not independent, i.e $a \not\perp b$. 
  We say that a graphical model $\sG$ is \emph{faithful}
  if for any variables $a,b,c$
  such that there is an active trail\footnote{See \todo{cite textbook} for a definition.} between $a$ and $b$ conditioned on $c$,
  then $a$ and $b$ are dependent given $c$ (i.e., $a\not\perp b \given c$).
\end{property}

%We will prove that the following simple property guarantees that all cliques
%have the exclusive views property:

%Applying \TensorFactorize to every bottleneck gives us a set of views
%  for every hidden variable. 
%To proceed, we need to check that \
%%(i) understand the assumptions under which
%  %\assumptionref{full-rank} holds for each bottleneck, and (ii)
%  that the candidate views actually imply that \propertyref{exclusive-views}
%  holds for every clique in the graph.

%Before we establish this relationship,
We now describe a reduction of any graphical model to
  a canonical form to make our arguments simpler.
  % DONE: don't say this yet since reader can't appreciate it.
  %that will allow us to handle the case where an observed
  %variable has multiple parents.

\begin{lemma}[Canonical form]
  \label{lem:reduction}
Every directed (undirected) graphical model can be transformed into one in which
  the observed variables are leaves with exactly one parent (neighbor). 
There is a one-to-one correspondence between the parameters of the
  transformed and original models.
\end{lemma}
\begin{proof}
  \begin{figure}
    \centering
    \subimport{figures/}{reduction.tikz}
    \caption{Reduction to canonical form.}
    \label{fig:reduction}
  \end{figure}

  \providecommand{\hp}{\ensuremath{h_\text{\rm new}}}
  Let $x_v$ be an observed variable with parents $\Pa(x_v)$ and children $\text{Ch}(x_v)$.
  Consider the following transformation.
  Replace $x_v$ with a new hidden variable \hp\ with the same
  parents $\Pa(x_v)$ and children $\text{Ch}(x_v) \union \{x_{v_1}, x_{v_2}, x_{v_3}\}$,
  where $x_{v_1},x_{v_2},x_{v_3}$ are three copies of $x_v$
  (\figureref{reduction}).
  Define $\Pr(\hp \mid \Pa(x_v)) = \Pr(x_v \mid \Pa(x_v))$ and
  $\Pr(x_{v_j} \mid \hp) = I$.
  %$\Pr(x_{v_j} \mid \hp) = \left[ \begin{array}{c} I_{k \times k} \\ 0_{(d-k) \times k} \end{array} \right]$.
  %The domain of $\hp$ is the same as $x_{v_1},x_{v_2},x_{v_3}$.
  % PL: cut to save space
  %\footnote{
      %Though we have assumed that all hidden variables share the
      %same domain, $[k]$, our work generalizes to hidden domains of
      %different sizes, provided \assumptionref{full-rank} is satisfied.
      %},
    %and $\mOpp{v}{\textrm{new}} = I$.
  Then, there is a one-to-one correspondence between every value of
  $\hp$ and $x_v$. Consequently, for any clique $\sC \contains \hp$, the
  parameters in the original graphical model can be obtained by
  substituting $\hp$ with $x_v$.

  This procedure applies straightforwardly for undirected graphical
  models, considering the neighbors $\sN(x_v)$ instead of its parents
  and children.
\end{proof}

Now we establish the link between bottlenecks and exclusive views
(see \appendixref{exclusive-views} for the complete proof):
\begin{lemma}[Uniformly bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
  %a connected subgraph of hidden variables.
  %If every hidden variable
  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
  If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
  $\sG$ has the exclusive views property.
\end{lemma}
\begin{proof}
We prove the contrapositive: suppose that there is some hidden clique $\sC \in \sG$ with $h_0 \in \sC$
that does not have exclusive views.  This means each observed variable $x_i$ is conditionally dependent on some
$h_i \in \sC \backslash \{h_0\}$ given $h_0$.
Consider any such two $x_1, x_2$.
Conditioned on $h_0$, there exists an active trail between them (specifically, $x_1 - h_1 - h_2 - x_2$).
Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
As this holds for any $x_1,x_2$, $h_0$ cannot be a bottleneck.
\end{proof}

@@

      \FOR{each hidden variable $h_i \in H$} 
        \STATE Apply $\TensorFactorize$ to learn conditional moments
        $\mOpp{v}{i}$ for every $v \in \sV_{h_i}$.

@@

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$. % (i.e. we have parameter sharing).
While the first and last hidden variables $h_1, h_M$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_M$, respectively)
  due to parameter sharing.
  %whose parameters we know because they share parameters, $O$.
%We first use the bottleneck $h_2$ with views $x_1, x_2, x_3$ to estimate $O$,
%and then recover $\pi$ from $\{h_1\}$ and recover $T$ from the clique $\{h_{1}, h_{2}\}$.

@@

In the latent tree model (\figureref{examples-tree}), the parameters
are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).
  %Note that while
  %$x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
  %$h_1$.

%In step 1, we recover the parameters $O$ from the bottleneck $h_2$ with
%  views $\{x^a_2, x^b_2, x^a_3\}$. We also recover the conditional moments
%  $\mOpp{2}{1}$, $\mOpp{3}{1}$, $\mOpp{4}{1}$ for $h_1$. 
%In step 2, we can recover $\pi$ from the clique $\{h_1\}$, using any
%  one of views (they are all exclusive). 
%To recover $T$ from the clique $\{h_1, h_2\}$, we use the views $x^a_2$
%  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_3$). Note that while
%  $x^a_3$ is also a view for $h_2$, $x^a_3$ is independent of $h_2$ given
%  $h_1$.

@@

%In a certain sense, having three views is necessary for identifiability;
%two views is certainly not enough.
% PL: perhaps too simple, saving space
%The simplest example of a model that cannot be recovered by our algorithm is
%  a mixture model with a single view, i.e. \figureref{three-view} with
%  just the variables $h_1$ and $x_1$.
%However, without further assumptions the model itself is not
%  identifiable\footnote{To see this, observe that probability-mass
%  can be arbitrarily exchanged between $\Pr(h_1)$ and $\Pr(x_1 | h_1)$ while preserving the marginals $\Pr(x_1)$}.
%A class of binary-valued noisy-or networks
%(see \figureref{examples-noisy-or} for an example)
%which do not possess the
%  bottleneck property are nonetheless identifiable;
%  see \citet{halpern13noisyor} for an algorithm.
%Having three views is in some sense necessary for the uniqueness of tensor factorization.
There are certainly models which are identifiable but do not have exclusive views.
For example, \figureref{examples-noisy-or} shows
  a binary-valued noisy-or network which can be
  learned by the algorithm of \citet{halpern13noisyor},
  but does not satisfy the exclusive views property.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Sample complexity}
\label{sec:sampleComplexity}
% PL: "combine" is too vague to be useful
%\LearnMarginals combines two consistent algorithms, \TensorFactorize and
%\LearnCliqueNs, and is thus consistent itself.

%In this section, we provide formal conditions under which $\LearnMarginals$
%will produce consistent estimates.
%We let
%$\mOpphat{v}{i}$,
%$\hat Z_\sC$,
%and $\hat M_\sV$,
%denote estimators for
%$\mOpp{v}{i}$,
%$Z_\sC$,
%and $M_\sV$,
%respectively.

% But $\mOpp{v}{i}$ is a product of tensors on the path from $h_i$ to
%   $x_v$, marginalizing out the rest of the graph.
% To make this property more interpretable, let us assume the following
%   property which we prove to be sufficient in
%   \appendixref{assumption-proof}:
% \begin{assumption} 
% \label{asm:full-rank-plus}
% The marginal distribution of every latent variable has full support:
%   $\forall{h_i \in H},~ \mPi{i} \succ 0$.
% For every clique $\sC \in \sG$ (including ones with observed variables),
%   and every conditional distribution $\Pr(h_c \given \Pa(h_c))$, where
%   $h_c \in \sC$, every {\em tensor unfolding} has full column rank, $k$.
% \end{assumption}
% In our directed grid example (\sectionref{directedExample}), for the
%   clique $\sC = \{ h_1, h_2 \}$ this condition implies $\pi = \Pr(h_1)
%   \succ 0$ and $T = \Pr(h_2 \mid h_1)$ has column rank $k$; for the clique
%   $\sC = \{ h_1, x_1 \}$, it implies $O = \Pr(x_1 \mid h_1)$ has column
%   rank $k$.

%\todo{did they prove this; if so just say "they show", not "we can show"}
%Using results from
%  \citet{anandkumar12moments,anandkumar13tensor} we can show that
%  learning $\mOpp{v_1}{i}$ for the bottleneck $h_i$ with views $x_{v_1},
%  x_{v_2}, x_{v_3}$ has the following sample complexity,
%  \todo{define hat O}
%  \todo{max subscript is misaligned}
%\begin{align*}
%  \|\mOpphat{v_1}{i} - \mOpp{v_1}{i}\|^2_F &= \frac{1}{\sqrt{n}} O\left( \frac{k {\pi\oft{i}}_{\max}^2}{\sigma_k(M_{v_1,v_2})^5} \right). 
%\end{align*}
It is easy to see that $\LearnMarginals$ has a polynomial sample complexity because it composes two parts that individually have polynomial sample complexity.
From \citet{anandkumar13tensor} given the non-degeneracy assumptions
  \assumptionref{non-degeneracy}, we have that $\mOpphat{v}{i}$
  converges to $\mOpp{v}{i}$ at a rate of $n^{-\frac12}$ with a constant
  that depends polynomially on the $k$-th singular value of
  $\mOpp{v}{i}$.
% PL: we don't have a crisp theorem (because there's multiple paths), so leave it vague.
Recovering the marginals $Z_\sC$ from the conditional moments via $\LearnClique$ is a linear operation and thus also has polynomial sample complexity. 
Note that $\sigma_{k}(\mOpp{v}{i})$ can become extremely
small if $h_i$ and $x_v$ are connected via many intermediate hidden variables.
\todo{cite work - Sontag?}
%\begin{theorem}[Sample complexity for $\mOpp{v}{i}$]
%  \label{thm:sample-complexity}
%  Without loss of generality, let $v=1, i=1$, and let $x_1, x_2, x_3$ be three views for
%  $h_1$. With probability at least $1 - \delta$,
%\begin{align*}
%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
%    &\le  \\
%    &
%  \hspace{-0.3in}
%      O\left(
%      % PL: remove pi_max because that's upper bounded by 1.
%      \frac{k \log(1/\delta) } 
%      {\sqrt{n} \, \pi\oft{1}_{\min} (\sigma_{k}(\mOpp{1}{1}) \sigma_{k}(\mOpp{1}{2}))^{5/2}} \right),
%\end{align*}
%where $\pi\oft{1}_{\min}$ is the smallest element of $\pi\oft{1}$.

%If $\sigma_{\min}(\Pr(c \given \Pa(h_c))) \le \sigma$ for every clique
%  $\sC \in \sG$ and hidden variable $h_c \in \sC$,
%\begin{align*}
%  \|\mOpphat{1}{1} - \mOpp{1}{1}\|_F 
%    &\le 
%      O\left( 
%      \frac{k \log(1/\delta) {\pi\oft{1}}_{\max}/{\pi\oft{1}}_{\min}} 
%      {\sqrt{n} \sigma^{5t}} \right) \epsilon,
%\end{align*}
%where $t$ is the length of the shortest topological ordering from $h_1$
%to the unique parent of $x_1$, $h_t$.
%\end{theorem}
%In general, $\sigma_{k}(\mOpp{v}{i})$ will depend exponentially on the
  %path length between the hidden variable $h_i$ and $x_v$.
%This is expected because the influence from $h_i$ to $x_v$
  %gets diluted for every latent variable it must factor through.

%\theoremref{sample-complexity}, proved in \appendixref{assumption-proof}
%  says that the number of samples required to estimate the view
%  $\mOpp{v}{i}$ grows exponentially in the {\em distance} between $x_v$
%  and $h_i$.

%\todo{Present sample complexity bounds for $Z_\sC$?}

>>>>

\section{Exclusive views}
\label{sec:directed}

%In this section, we will develop a consistent parameter estimate for a class of directed graphical models.
%The conditional moments $\mOpp{v}{i} \eqdef \Pr(x_v \mid h_i)$
%recovered from {\TensorFactorize}, are not the underlying parameters.
  %Once we recover the conditional moments $\mOpp{v}{j} \eqdef \Pr(x_v
  %| h_j)$ for some $v, j$, we present a systematic approach to learn the
  %conditional probability tables for every clique for a class of directed
  %graphical models.
In this section, we will use the conditional moments $\mOpp{v}{i} \eqdef
\Pr(x_v \mid h_i)$ recovered from the previous section
to compute the marginal distribution of sets of hidden variables $Z_S \eqdef \Pr(\bh_S)$,
taking us one step closer to the actual model parameters.
For clarity, we derive our algorithm in the infinite data limit,
although the actual algorithm works with estimates.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Example: directed grid model}
\label{sec:directedExample}

To gain some intuition, consider the directed grid model from \figureref{approach}.
%The model has eight observed variables $x^a_1, x^b_1 \cdots, x^a_4, x^b_4$ and four
  %hidden variables $h_1, \ldots, h_4$.
The parameters of this model are the conditional probability tables
$\pi \eqdef \Pr(h_1) \in \Re^k, T \eqdef \Pr(h_2 | h_1) = \Pr(h_3 | h_1) \in \Re^{k \times k},
V \eqdef \Pr(h_4 | h_2, h_3) \in \Re^{k \times k \times k}$ and $O \eqdef \Pr(x^a_i | h_i)
=  \Pr(x^b_i | h_i) \in \Re^{d \times k}$.

\paragraph{Estimating $O$}
Since the observed variables $x^a_1, x^b_1, x^a_2$ are
  conditionally independent given $h_1$, we can $\TensorFactorize$ from
  \sectionref{setup} to recover $O$.

\paragraph{Estimating $\pi$}
The moments of $x^a_1$, $\mO_1 \eqdef \Pr(x^a_1)$ are directly related to
  $\pi$ by a linear system: $\mO_1 = O \pi$. 
If $O$ has full column rank, we can recover $\pi$ by taking the pseudoinverse: $\pi = O\pinv  \mO_1$.

@@

  %Section~\ref{sec:directedGeneral} will describe the algorithm in full generality.
%For ease of exposition we make the following simplifications to our presentation:
%(i) we describe our algorithm entirely in the context of directed
  %graphical models, though it generalizes to undirected models;
%(iii) we present the algorithm solely in terms of linear algebraic operations;
  %\sectionref{piecewise} provides a statistically
  %more efficient estimator using composite likelihood.

The intuitions of the directed grid generalize readily to general graphs.
The key property required for our algorithm to work is as follows:
\begin{property}[Exclusive views]
  \label{prop:exclusive-views}
Given a graphical model $\sG$,
  a subset of hidden variables $S \subset \bh$ is said to satisfy the \emph{exclusive views property} if for
  each $h_i \in S$, the two conditions hold:
  (i) there exists some observed variable
  $x_{v}$ which is conditionally independent of the others $S \backslash \{ h_i \}$ given $h_i$,
  and (ii) the conditional moment matrix $\mOpp{v}{i} \eqdef
  \Pr(x_{v} \mid h_i)$ have full column rank $k$ and can be recovered.
We call $x_v$ an exclusive view of $h_i$ in $S$.
%If every hidden clique has the exclusive views property, then we say that $\sG$ does as well.
\end{property}

@@

We now show that if a subset of hidden variables $S$ has the exclusive views property,
then we can recover the marginal distribution $\Pr(\bh_S)$.
Consider any $S = \{i_1, \ldots, i_m\}$ with the exclusive views property. Let
  $x_{v_j}$ be an exclusive view for $h_{i_j}$ in $S$ and define $\sV
  = \{v_1, \ldots, v_m\}$. % be a set of exclusive views for the clique $S$.
By the exclusive views property,
the marginal over the observed variables $\Pr(\bx_\sV)$
factorizes according to the marginal over the hidden variables $\Pr(\bh_S)$
times the conditional moments:
\begin{align*}
  \mO_\sV 
  &\eqdef \Pr(\Sx{\sV}) \\
  &= \sum_{\vh_S} \Pr(\Sh{S}) 
                    \Pr(x_{v_1} | h_{i_1}) \cdots \Pr(x_{v_m} | h_{i_m}) \\
  &= Z_{S}(\mOpp{v_1}{i_1},\dots,\mOpp{v_m}{i_m}) \\
  &= Z_{S}(\mOppAll),
\end{align*}
where $\mOppAll = \mOpp{v_1}{i_1} \otimes \cdots \otimes \mOpp{v_m}{i_m}$ is the tensor product of
all the conditional moments.
Vectorizing, we have that
$Z_S \in \Re^{k^m}$,
$M_\sV \in \Re^{d^m}$,
and $\mOppAll \in \Re^{d^m \times k^m}$.
Since each $\mOpp{v}{i}$ has full column rank $k$,
$\mOppAll$ has full column rank $k^m$.
Succinctly, $\mO_\sV$ (which can be estimated directly from data)
is a linear function of $Z_S$ (what we seek to recover).
We can solve for the hidden marginals $Z_S$ simply by projecting $\mO_\sV$ onto the column
space of $\mOppAll$:
\begin{align*}
  Z_{S} &= \mO_\sV(\mOpp{v_1}{i_1}\pinv,\cdots,\mOpp{v_m}{i_m}\pinv).
\end{align*}

\algorithmref{learnclique} summarizes the procedure, \LearnClique.
Given $Z_S$, the conditional probability tables for $S$ can easily be
obtained via renormalization.
\begin{theorem}[Hidden marginals from exclusive views]
If $S \subseteq \bx$ is a subset of hidden variables with the exclusive views property,
then \algorithmref{learnclique} recovers the marginals $Z_S = \Pr(\bh_S)$.
\end{theorem}

%If each $\mOpp{v_j}{i_j}$ has full column rank, then

\begin{algorithm}
  \caption{\LearnClique~(pseudoinverse version)}
  \label{algo:learnclique}
  \begin{algorithmic}
    \REQUIRE Hidden subset $S$ with exclusive views (\propertyref{exclusive-views}).
    \ENSURE Marginal distribution $Z_S$.
      \STATE Identify exclusive views $\Sx{\sV} = \{x_{v_1}, \dots, x_{v_m}\}$.
      \STATE Return $\mH_S \gets \mO_{\sV}( \mOpp{v_1}{i_i}\pinv, \dots, {\mOpp{v_m}{i_m}}\pinv )$.
  \end{algorithmic}
\end{algorithm}

@@

We have shown how to estimate the parameters of any clique that possesses the exclusive
  views property (\propertyref{exclusive-views}).
  But in a general graph, which cliques have this property?
  To provide intuition, we will provide more interpretable sufficient conditions
  and examples.

It is clear that having uniform bottlenecks implies part (ii)
of the exclusive views property (\propertyref{exclusive-views}) via \TensorFactorize,
but we will show that part (i) follows as well.

The exclusive views property 
Next, we define a (in analogy with biconnected component)
\begin{definition}[Bidependent set]
Given a graphical model $\sG$, we say that a subset of nodes $S$ is \emph{bidependent} if
conditioned on any $a \in S$, there is an active trail between any other two nodes $b,c \in S$.
\end{definition}
Intuitively
Note that all cliques are bidependent, but 
Bottlenecks provide three views, but to satisfy condition (ii) of \propertyref{exclusive-views},
only two is required.

Now we establish the link between bottlenecks and exclusive views:
\begin{lemma}[Uniformly bottlenecked implies exclusive views]
  \label{lem:bottleneck-views}  
% PL: simplify to cliques (this is also bad notation, since \sG is a set of cliques, not nodes)
%Let $\sG' = \{ h_{i_1}, \dots, h_{i_m} \} \subseteq \sG$ be
  %a connected subgraph of hidden variables.
  %If every hidden variable
  %$h_i \in \sG'$ is a bottleneck, then $\sG'$ has the exclusive views property.
  If a graphical model $\sG$ is faithful and uniformly bottlenecked, then
  every bidependent set in $\sG$ has the exclusive views property (and hence so does $\sG$).
\end{lemma}
\begin{proof}
We prove the contrapositive: suppose there exists a biconnected subgraph of $\sH \subset \sG$ consisting of hidden nodes
such that some $h_0 \in \sH$ does not have an exclusive view.
This means each observed variable $x_i$ in the graph is conditionally dependent on some
$h_i \in \sH \backslash \{h_0\}$ given $h_0$.
Consider any such two $x_1, x_2$.
Conditioned on $h_0$,
there is an active trail between $h_1$ and $h_2$ because $\sH$ is biconnected.
Thus, there is an active trail $x_1 - h_1 - h_2 - x_2$ conditioned on $h_0$.
Since $\sG$ is faithful, we have $x_1 \not\perp x_2 \mid h_0$.
As this holds for any $x_1,x_2$, $h_0$ cannot be a bottleneck.
\end{proof}

@@

      \FOR{each hidden variable $h_i \in \bh$} 
        \STATE Apply $\TensorFactorize$ to learn conditional moments
        $\mOpp{v}{i}$ for every $v \in \sV_{h_i}$.

@@

In the HMM (\figureref{examples-hmm}), the parameters
are $O \eqdef \Pr(x_i|h_i)$ and $T \eqdef \Pr(h_{i+1} | h_i)$
for all $i$.
While the first and last hidden variables $h_1, h_M$ in the
  sequence are not bottlenecks, they still have exclusive views ($x_1$ and
  $x_M$, respectively)
  due to parameter sharing.

@@

In the latent tree model (\figureref{examples-tree}), the parameters
are $\pi \eqdef \Pr(h_i)$, $T \eqdef \Pr(h_i | h_1)$ for $i \in \{2,3,4\}$,
and $O \eqdef \Pr(x^a_i | h_i) = \Pr(x^b_i | h_i)$ for $i \in \{2,3,4\}$.
Note that while $h_1$ is not directly connected to an observed variable,
  it is still a bottleneck, with views $x^a_2, x^a_3, x^a_4$.
We can recover $T$ from the clique $\{h_1, h_2\}$ by using views $x^a_2$
  (exclusive to $h_2$) and $x^a_3$ (exclusive to $h_1$).

@@

There are certainly models which are identifiable but do not have exclusive views.
For example, \figureref{examples-noisy-or} shows
  a binary-valued noisy-or network which can be
  learned by the algorithm of \citet{halpern13noisyor},
  but does not satisfy the exclusive views property.
